\documentclass[11pt]{article}
\usepackage{agda}
\usepackage{fontspec}
\usepackage{newunicodechar}
\usepackage{etoolbox}
\setsansfont{Arial Unicode MS}
\usepackage[margin=1in]{geometry}
\usepackage[hidelinks]{hyperref}
\newunicodechar{φ}{\ensuremath{\phi}}

% Match the main Void and Form code layout: compact Agda blocks,
% no extra math indentation, and narrower indentation steps.
\renewcommand{\AgdaCodeStyle}{\fontsize{8}{9}\selectfont}
\setlength{\mathindent}{0pt}
\renewcommand{\AgdaIndentSpace}{\hspace*{0.30em}}
\AtBeginEnvironment{code}{\hfuzz=2pt}

\title{Void Companion: Binary Distinction and the Four-Vertex Closure}
\author{Johannes Michael Wielsch\\
\small Companion note to \emph{Void and Form}\\
\small Project DOI: \href{https://doi.org/10.5281/zenodo.19491035}{10.5281/zenodo.19491035}}
\date{May 2026}

\begin{document}
\maketitle

\begin{abstract}
This note is the compact map and self-contained formal kernel of
\emph{Void and Form}.  It does not ask what emerges from a presupposed
``nothing''.  It asks under which conditions formal description is
possible at all.  The pre-formal answer is the availability of
distinguishable positions: without them no relation, judgement,
equation, map, representation, truth-value, or falsehood can be
formulated.  \emph{Void} names this boundary.  It is not the empty set,
not the empty type, and not a physical nothing; it is the absence of the
domain in which those notions could already occur.

The checked kernel begins once distinguishability is granted as data.
In intensional Martin-L\"of type theory, a binary distinction is
represented by two non-collapsing points that exhaust their carrier.
This carrier has canonical normal form \AgdaDatatype{Two}; its
endomorphisms exhaust to four cases; representing those cases with
separation, complete realisation, and no surplus yields a four-vertex
carrier with complete simple adjacency; and every such closure points
back to the data from which it was built.  In addition, the cardinality
of the entry datum is itself made formal: a singleton carrier admits no
\AgdaRecord{Distinction}; the original record is exactly the $n=2$ case
of an n-ary template; and at $n=3$ the n-ary record is definitionally
equivalent to a binary package of three pairwise separations and a
nested binary cover.

The note does not claim that such a distinction takes place.  It
records what follows when the formal threshold is crossed.  The thesis
that distinction is foundational is at least as old as Spencer-Brown's
\emph{Laws of Form} and is not asserted here as new.  The contribution
is the precise, machine-checked statement that, under the explicitly
fixed conditions of representation, the closure of binary distinction
is $K_4$, unique up to isomorphism, and dependent back on the data from
which it was built.
\end{abstract}

\section*{How to read this note}

This is the self-contained formal kernel of \emph{Void and Form}, not
the full development.  It can be read before the larger source.  For
readability, it uses standard-library imports for equality, sums,
products, the empty type, and negation.  The full file
\texttt{Void.lagda.tex} rebuilds that infrastructure internally and
continues with constructions that are out of scope for this note.

The note has two layers.  First, it names the pre-formal boundary at
which formalisation becomes possible at all: relations, morphisms,
judgements, and equations already presuppose a field in which one place
can be separated from another.  This is the conceptual reason for
taking binary distinction as the starting datum; it is not an Agda
theorem.  Second, it records the checked statements: normal form,
exhaustion, four-vertex closure, and dependency back to the starting
data.

The code is narrower.  It starts only after the data are granted.  From
that point on, every formal claim below is a term checked by Agda.  The
tetrahedron is only the geometric reading of the complete graph on the
four endomorphism cases; it is not an additional assumption used in the
proof.

\section*{Map of the project}

This note is also the table of contents for the project in compressed
form.  The map below is not a second proof.  It fixes the status of the
layers so that the reader can see where \emph{Void} ends, where the
formal threshold begins, and where \emph{Form} later becomes
interpretation.

\subsection*{Void: conditions of formulability}

\begin{description}
\item[Level 0: Void.] No type, no term, no carrier.  The schematic
marker is $\mathcal{V} \notin \mathit{Set}_\omega$.  Void is not a
stable endpoint in the sense of a foundation, because every act of
positing already presupposes structure.  It is therefore conceptually
non-closable.  This level is marked, not proved.

\item[Level 1: Information.] Information is understood in Bateson's
minimal sense: a difference that makes a difference.  Without
distinction, an occurrence carries no information.  Information
therefore presupposes distinction, and an inhabited distinction
requires at least two distinguishable occurrences.  This binary shape
is minimal because every further difference already relies on it.

\item[Level 2: Two.] The first stable formal carrier is
\AgdaDatatype{Two}.  It contains two separated points and no third
generator.  It is not a truth-value object; it is the normal form of
distinction.  From this level onward the formal threshold is crossed:
MLTT and Agda can execute the consequences of the datum they have
received.

\item[Level 3: Logic.] Formal systems operate on structures in which
distinguishable positions are already available.  Relation, morphism,
judgement, and equation do not construct that availability; they use
it.  MLTT is the clean carrier used here, not the origin of the cut.

\item[Level 4: Mathematics.] The machine-checked chain is the
mathematical core: every \AgdaRecord{Distinction} is isomorphic to
\AgdaDatatype{Two}; the endomorphisms of \AgdaDatatype{Two} have exactly
four cases; representing those cases under separation, complete
realisation, and no surplus forces the four-element closure, unique up
to isomorphism; and the closure returns the originating distinction.
No further generator is introduced.
\end{description}

\subsection*{Form: structural reading and interpretation}

\begin{description}
\item[Level 5: Structural separation.] Two distinguishable occurrences
collapse if neither origin nor position can distinguish them.
Non-identity of origin gives the germ of temporal order;
non-coincidence gives the germ of spatial separation.  Space and time
are not asserted as primitive entities here.  They name conditions
under which distinction does not collapse.

\item[Level 6: Physics.] Physics enters as interpretation of the
preceding structural conditions by an observer who already operates
inside them.  Form is not physics; it is the structure from which
physical description can be read.

\item[Level 7: Contingency.] Necessity stops before complete worldly
specification.  The preceding levels fix conditions, not the full
concrete world that satisfies them.

\item[Level 8: Induction.] No new external generator is postulated.
More complex structures are read as repeated applications and
stabilisations of the original distinction.  Induction is the law of
that repetition.  This is a structural reading, not a reductionist
claim about all mathematics.
\end{description}

Levels 0 and 1 are not formal statements inside a type system.  They
state the conditions under which formal statements can appear.  From
Level 2 onward the formal theory begins, and the claims are expressible
and checkable inside it.

\part*{Part I.\quad The pre-formal threshold}

\section*{The pre-formal boundary}

The project begins before the point at which a formal theory can already
speak in its own voice.  A relation has two positions.  A morphism has a
source and a target.  A judgement has a term and a type.  An equation has
two sides, even when it proves them equal.  Every such form is already a
use of separation.  A theory may analyse that separation once it has been
given; it cannot produce the very field of separability from within a
field that it already presupposes.

The following display is therefore not Agda code and not a theorem of
MLTT.  It is a schematic boundary marker: it says where ontology hands
the first stable datum to formalisation.  The names in the display are
not internal definitions, and the arrows are not checked implications;
they are meta-level dependency markers for the passage into the formal
part.

\[
\begin{array}{ll}
\mathrm{Level\;0:} &
  \mathcal{V} \notin \mathit{Set}_\omega,
  \qquad \neg\exists x\,(x : \mathcal{V}).
\\[0.6em]
\mathrm{Level\;1:} &
  \mathrm{Sing}(S) :=
  \exists p:S\,\forall x:S\,(x = p).
\\[0.6em]
\mathrm{Collapse:} &
  \mathrm{Sing}(S)
  \Longrightarrow
  \neg\exists a,b:S\,(a \neq b).
\\[0.6em]
\mathrm{Level\;2:} &
  \mathrm{Dist}(S) :=
  \exists a,b:S\,
  \bigl(a \neq b\ \wedge\ \forall x:S\,(x=a \vee x=b)\bigr).
\\[0.6em]
\mathrm{Separation:} &
  a \neq b
  \Longrightarrow
  \neg\mathrm{SameOrigin}(a,b)\ \wedge\
  \neg\mathrm{Contact}(a,b).
\\[0.6em]
\mathrm{Formal\;threshold:} &
  \mathrm{Dist}(S)
  \Longrightarrow
  \begin{array}[t]{l}
  S \simeq \mathbf{2},\\
  \mathrm{End}(\mathbf{2}) =
    \{c_L,c_R,\mathrm{id},\sigma\},\\
  \mathrm{representation\ closure} = K_4,\\
  K_4 \Longrightarrow \mathrm{Dist}(S).
  \end{array}
\end{array}
\]

Level 0 is not a hidden type.  The symbol \(\mathcal{V}\) marks the
absence of a domain in which a term could already be typed.  Level 1
shows why mere inhabitance is still collapse: if all occurrences are
the same occurrence, no difference can be carried internally.  Level 2
is the first non-collapsing datum: two separated occurrences and no
third generator.  Here, \(\mathrm{SameOrigin}(a,b)\) names the collapse
in which the two occurrences have no distinguishable order of arising;
\(\mathrm{Contact}(a,b)\) names the collapse in which no gap separates
their positions.  The two negations in the separation line mark the
ontological content that the later volume \emph{Form} reads as the
germ of temporal and spatial structure: without an interval of origin
or a gap of contact, the two occurrences collapse back into one.

This boundary is not a theorem of the development; it is the condition
under which the development can be read as a theorem at all.
Only the last line enters Agda.  From \(\mathrm{Dist}(S)\) onward, the
schematic notation is replaced by checked terms.  The rest of this note
records that checked consequence.

\section*{Why two cannot occur without separation}

Before any cardinality is even chosen, a sharper question arises.  Why
can two occurrences not already be present at the threshold without
any separating relation between them?  In the absence of time, space,
or any other field of differentiation, why does the supposed pair not
simply coexist as two?

The answer is not specific to type theory and not specific to any
single foundation.  It is the principle that what cannot be told apart
is not two.  Leibniz stated it in 1686 as the
\emph{principium identitatis indiscernibilium}: indiscernibles are
identical.  The same content reappears, in the technical idiom of each
framework, wherever ``two'' is given a precise sense:

\begin{itemize}
\item In classical first-order logic with equality, substitutivity of
  identity makes any two elements that satisfy the same predicates
  equal.
\item In set theory, the axiom of extensionality says that two sets
  with the same elements are the same set.
\item In category theory, the Yoneda principle expresses that an
  object is determined by its relations to others; an object with no
  distinguishing relation is not a second object.
\item In quantum mechanics, particles in the same state are not two
  entities that happen to coincide; the statistical structure of the
  theory (Bose, Fermi) is forced precisely because indistinguishables
  cannot be counted as two.
\end{itemize}

Each of these frameworks pays the same price: as soon as ``two
distinct $X$'' is even said, separation has already been used.  No
foundation has exhibited a coherent counterexample, because exhibition
itself requires a distinguishing predicate.

The consequence is immediate.  Whatever has no separating predicate
is, by the principle that defines what equality of occurrences means,
one occurrence and not two.  The first non-collapsing pair and the
first separating predicate arrive together; there is no temporal,
spatial, or relational gap in which two might quietly coexist before
separation.  The absence of any separating predicate forces collapse
to one.

The companion does not assume separation as a primitive act in the
manner of \emph{Laws of Form}; it relies on a principle that any
framework supporting the very statement ``two distinct $X$'' already
concedes.  How this principle becomes a checkable term once a specific
carrier is chosen is shown later, in the section on the executed
Leibniz collapse, after the formal infrastructure of this note is in
place.

\section*{Why binary, not unary or ternary}

The choice of binary distinction as the entry datum is not stipulation
and not preference.  At the pre-formal threshold, only one cardinality
survives the joint demand of non-collapse and self-sufficiency.  This
argument is conceptual, not an Agda theorem; the formal results below
do not depend on it.  It is recorded here so that the antecedent of
the machine-checked conditional is not itself opaque.

\emph{Unary collapses.}  A single occurrence carries no separation.
The schematic predicate $\mathrm{Sing}(S)$ records exactly this: every
element equals the chosen point, and no pair $(a,b)$ with $a\neq b$
can be exhibited.  Whatever appears within $S$ is the same appearance.
Formalisation cannot start here, because relations, morphisms,
judgements, and equations all require two positions before they can be
written down.

\emph{Ternary carries derived binary plus surplus.}  Three positions
$a,b,c$ require three pairwise separations $a\neq b$, $b\neq c$,
$a\neq c$.  Each pairwise separation is itself a binary distinction.
A primitive ternary datum would have to supply, on top of these three
binary facts, an additional structure (an order, an incidence, a
selection of one position against the other two) that the act of
distinguishing alone does not yield.  Any device that inspects three
positions simultaneously is itself a fourth position relative to the
three, and its application to them again splits as ``this one against
the other two''.  The ternary case therefore decomposes into binary
distinctions plus surplus structure that requires its own justification.

\emph{Binary survives.}  A binary distinction supplies exactly one
separation $a\neq b$ together with the exhaustion clause that nothing
further is present.  No additional structure (order, incidence,
selection) is assumed; any such structure that arises later is forced
by the four endomorphism cases of the normal form, not added to the
datum.  This is why the formal development below begins with
\AgdaRecord{Distinction} and not with a richer or poorer record.

These three observations are made formal in Part~II, in the section
on the cardinality of distinction: a singleton carrier admits no
inhabitant of \AgdaRecord{Distinction}, the n-ary version of the
record is at $n=2$ mutually convertible with \AgdaRecord{Distinction},
and higher arities carry only pairwise binary separation witnesses
together with a nested binary disjunction.

\part*{Part II.\quad The formal threshold}

\section{Setting}

The setting is ordinary intensional Martin-Löf type theory as
implemented in Agda.  The code snippets below use the standard library
for equality, sums, products, the empty type, and negation.  No attempt
is made here to reproduce the full formal development; the goal is to
record the minimal spine of the argument.  The carrier of the binary
distinction is introduced as a fresh type \AgdaDatatype{Two} with
constructors \AgdaInductiveConstructor{L} and
\AgdaInductiveConstructor{R}, rather than reusing \AgdaDatatype{Bool}.
The two types have the same shape, but \AgdaDatatype{Bool} carries an
established reading as truth-values.  The result below depends on
separation alone, not on any prior reading of the two points as
truth or falsity, and the renaming makes this independence explicit.

\begin{code}
{-# OPTIONS --safe --without-K #-}

module VoidCompanion where

open import Relation.Binary.PropositionalEquality using (_≡_; refl; sym; trans; cong)
open import Data.Empty using (⊥; ⊥-elim)
open import Data.Sum using (_⊎_; inj₁; inj₂)
open import Data.Product using (Σ; _,_)
open import Relation.Nullary using (¬_)
\end{code}

The two-point normal form is the following type.

\begin{code}
data Two : Set where
  L R : Two

L≠R : ¬ (L ≡ R)
L≠R ()

R≠L : ¬ (R ≡ L)
R≠L p = L≠R (sym p)
\end{code}

\section{Distinction}

A binary distinction is not just a type with two named points.  The
points must be separated, and every element of the carrier must be one
of them.  This is the exact hypothesis that makes the normal-form
theorem true.

\begin{code}
record Distinction : Set₁ where
  field
    S         : Set
    left      : S
    right     : S
    separated : ¬ (left ≡ right)
    cover     : (x : S) → (x ≡ left) ⊎ (x ≡ right)
\end{code}

The canonical inhabitant is obtained from \AgdaDatatype{Two} itself.

\begin{code}
Two-cover : (x : Two) → (x ≡ L) ⊎ (x ≡ R)
Two-cover L = inj₁ refl
Two-cover R = inj₂ refl

Two-distinction : Distinction
Two-distinction = record
  { S         = Two
  ; left      = L
  ; right     = R
  ; separated = L≠R
  ; cover     = Two-cover
  }
\end{code}

\section{Cardinality of distinction}

The earlier prose in Part~I argued that one occurrence collapses,
three occurrences decompose, and two survives.  The argument was
explicitly marked as conceptual.  At this point, after the
\AgdaRecord{Distinction} record is in place, the same content can be
stated and proved formally.  Two facts together turn the binary
choice from preference into a theorem of the present setting: a
singleton carrier admits no inhabitant of \AgdaRecord{Distinction} at
all, and the n-ary generalisation of the record is, at $n=2$, mutually
convertible with the binary case.  Higher arities carry only the
pairwise binary separation witnesses one already has, plus an
exhaustive cover that is structurally a nested binary disjunction.
No primitive ternary or higher-arity operator is required.

\begin{code}
open import Data.Nat using (ℕ; zero; suc)
open import Data.Fin using (Fin) renaming (zero to fz; suc to fs)

record Distinction-n (n : ℕ) : Set₁ where
  field
    S        : Set
    point    : Fin n → S
    distinct : (i j : Fin n) → ¬ (i ≡ j) → ¬ (point i ≡ point j)
    cover    : (x : S) → Σ (Fin n) (λ i → x ≡ point i)
\end{code}

\emph{Unary collapses formally.}  No singleton carrier hosts a
\AgdaRecord{Distinction}: if every two elements of the carrier are
equal, the boundary points are equal and \AgdaField{separated} is
falsified.

\begin{code}
singleton-no-distinction :
  (d : Distinction) →
  ((x y : Distinction.S d) → x ≡ y) →
  ⊥
singleton-no-distinction d hyp =
  Distinction.separated d (hyp (Distinction.left d) (Distinction.right d))
\end{code}

The same content holds for \AgdaRecord{Distinction-n} at $n=1$: any two
elements of its carrier are forced equal, so no separation between
distinct positions can be inhabited.

\begin{code}
distinction-n-1-collapses :
  (d : Distinction-n 1) →
  (x y : Distinction-n.S d) → x ≡ y
distinction-n-1-collapses d x y
  with Distinction-n.cover d x | Distinction-n.cover d y
... | fz , px | fz , py = trans px (sym py)
... | fz , _  | fs () , _
... | fs () , _ | _
\end{code}

The collapse on \AgdaRecord{Distinction-n} \AgdaNumber{1} is one half of
the ``unary kills distinction'' content; on its own, it forces every
two elements of the carrier to coincide.  The bridge to the original
record is the next step: any \AgdaRecord{Distinction} that injects
into such a carrier is immediately impossible, because the two
separated boundary points are forced equal under the injection and
\AgdaField{separated} is contradicted.  This makes the joint
statement, that no \AgdaRecord{Distinction} survives on a unary
carrier, a single checked term and not a verbal combination of two
lemmas.

\begin{code}
no-distinction-on-Dn1 :
  (d1 : Distinction-n 1) →
  (d  : Distinction) →
  (φ  : Distinction.S d → Distinction-n.S d1) →
  ((x y : Distinction.S d) → φ x ≡ φ y → x ≡ y) →
  ⊥
no-distinction-on-Dn1 d1 d φ φ-inj =
  Distinction.separated d
    (φ-inj (Distinction.left d) (Distinction.right d)
      (distinction-n-1-collapses d1
        (φ (Distinction.left d))
        (φ (Distinction.right d))))
\end{code}

\emph{Binary equals the existing record.}  At $n=2$ the n-ary record
is mutually convertible with the original \AgdaRecord{Distinction}.
The two constructions are mechanical and demonstrate that no
information is lost in either direction.

\begin{code}
distinction-to-Dn2 : Distinction → Distinction-n 2
distinction-to-Dn2 d = record
  { S        = Distinction.S d
  ; point    = pt
  ; distinct = dist
  ; cover    = cov
  }
  where
    open Distinction d
    pt : Fin 2 → S
    pt fz      = left
    pt (fs fz) = right

    dist : (i j : Fin 2) → ¬ (i ≡ j) → ¬ (pt i ≡ pt j)
    dist fz      fz      h _ = h refl
    dist fz      (fs fz) _ p = separated p
    dist (fs fz) fz      _ p = separated (sym p)
    dist (fs fz) (fs fz) h _ = h refl

    cov : (x : S) → Σ (Fin 2) (λ i → x ≡ pt i)
    cov x with cover x
    ... | inj₁ p = fz , p
    ... | inj₂ p = fs fz , p

Dn2-to-distinction : Distinction-n 2 → Distinction
Dn2-to-distinction d = record
  { S         = Distinction-n.S d
  ; left      = Distinction-n.point d fz
  ; right     = Distinction-n.point d (fs fz)
  ; separated = Distinction-n.distinct d fz (fs fz) λ ()
  ; cover     = cov
  }
  where
    open Distinction-n d
    cov : (x : S) → (x ≡ point fz) ⊎ (x ≡ point (fs fz))
    cov x with cover x
    ... | fz , p    = inj₁ p
    ... | fs fz , p = inj₂ p
\end{code}

\emph{Higher arity carries only pairwise binary plus a nested
disjunction.}  At any $n$, the n-ary distinction record carries a
binary separation witness for every ordered pair of distinct
positions.  This witness is the \AgdaField{distinct} field itself; no
further structure is needed to extract it.

\begin{code}
pairwise-binary :
  {n : ℕ} (d : Distinction-n n) →
  (i j : Fin n) → ¬ (i ≡ j) →
  ¬ (Distinction-n.point d i ≡ Distinction-n.point d j)
pairwise-binary = Distinction-n.distinct
\end{code}

At $n=3$, the exhaustive cover of three positions is structurally the
nested binary disjunction $A_0 \uplus (A_1 \uplus A_2)$, where $A_i$
records that an element equals position $i$.  No primitive ternary
disjunction is invoked; the three-case split is two binary splits.

\begin{code}
cover-3-nested :
  (d : Distinction-n 3) (x : Distinction-n.S d) →
  (x ≡ Distinction-n.point d fz)
    ⊎ ((x ≡ Distinction-n.point d (fs fz))
       ⊎ (x ≡ Distinction-n.point d (fs (fs fz))))
cover-3-nested d x with Distinction-n.cover d x
... | fz , p          = inj₁ p
... | fs fz , p       = inj₂ (inj₁ p)
... | fs (fs fz) , p  = inj₂ (inj₂ p)
\end{code}

The two preceding lemmas show only that a ternary record \emph{carries}
binary content.  The stronger claim, that the ternary record is
\emph{nothing more} than this binary content, requires the converse
direction: any package of three pairwise separations together with a
binary-nested cover must reconstruct a \AgdaRecord{Distinction-n}
\AgdaNumber{3}.  The following record names exactly the binary data
extractable from a ternary distinction, and the two conversions below
show the reduction in both directions.

\begin{code}
record Dn3-binary-data : Set₁ where
  field
    S     : Set
    p0    : S
    p1    : S
    p2    : S
    sep01 : ¬ (p0 ≡ p1)
    sep02 : ¬ (p0 ≡ p2)
    sep12 : ¬ (p1 ≡ p2)
    cover : (x : S) → (x ≡ p0) ⊎ ((x ≡ p1) ⊎ (x ≡ p2))

Dn3-to-binary : Distinction-n 3 → Dn3-binary-data
Dn3-to-binary d = record
  { S     = Distinction-n.S d
  ; p0    = Distinction-n.point d fz
  ; p1    = Distinction-n.point d (fs fz)
  ; p2    = Distinction-n.point d (fs (fs fz))
  ; sep01 = Distinction-n.distinct d fz (fs fz) λ ()
  ; sep02 = Distinction-n.distinct d fz (fs (fs fz)) λ ()
  ; sep12 = Distinction-n.distinct d (fs fz) (fs (fs fz)) λ ()
  ; cover = cover-3-nested d
  }

binary-to-Dn3 : Dn3-binary-data → Distinction-n 3
binary-to-Dn3 b = record
  { S        = S
  ; point    = pt
  ; distinct = dist
  ; cover    = cov
  }
  where
    open Dn3-binary-data b
    pt : Fin 3 → S
    pt fz             = p0
    pt (fs fz)        = p1
    pt (fs (fs fz))   = p2

    dist : (i j : Fin 3) → ¬ (i ≡ j) → ¬ (pt i ≡ pt j)
    dist fz             fz             h _ = h refl
    dist fz             (fs fz)        _ p = sep01 p
    dist fz             (fs (fs fz))   _ p = sep02 p
    dist (fs fz)        fz             _ p = sep01 (sym p)
    dist (fs fz)        (fs fz)        h _ = h refl
    dist (fs fz)        (fs (fs fz))   _ p = sep12 p
    dist (fs (fs fz))   fz             _ p = sep02 (sym p)
    dist (fs (fs fz))   (fs fz)        _ p = sep12 (sym p)
    dist (fs (fs fz))   (fs (fs fz))   h _ = h refl

    cov : (x : S) → Σ (Fin 3) (λ i → x ≡ pt i)
    cov x with cover x
    ... | inj₁ p          = fz , p
    ... | inj₂ (inj₁ p)   = fs fz , p
    ... | inj₂ (inj₂ p)   = fs (fs fz) , p
\end{code}

What these proofs together establish is the formal counterpart of
\emph{Why binary, not unary or ternary} from Part~I, in the precise
sense available to the present framework.  Singleton carriers cannot
host a distinction at all: the lemma
\AgdaFunction{distinction-n-1-collapses} forces every two elements of
a unary carrier to coincide, and \AgdaFunction{no-distinction-on-Dn1}
closes the bridge by showing that no \AgdaRecord{Distinction} can
inject into such a carrier without contradicting
\AgdaField{separated}.  The original record is exactly the $n=2$
case of an n-ary template (\AgdaFunction{distinction-to-Dn2},
\AgdaFunction{Dn2-to-distinction}).  At $n=3$, the conversions
\AgdaFunction{Dn3-to-binary} and \AgdaFunction{binary-to-Dn3} prove
more than that ternary data carries binary content: they prove that
ternary data is recoverable from binary content, since the binary
package of three pairwise separations and a nested binary cover
reconstructs a \AgdaRecord{Distinction-n} \AgdaNumber{3}.

The scope of this last claim is precisely the scope of the framework.
The theorem says that, inside this development, every inhabitant of
\AgdaRecord{Distinction-n} \AgdaNumber{3} is definitionally equivalent
to a binary package; it does not assert that no other foundation could
treat ``three positions'' as a primitive entity.  An ontology that
posits an irreducible ternary primitive is not refuted here.  What is
shown is that within the present record-based, --safe --without-K
fragment of MLTT, the n-ary record at $n=3$ adds no information beyond
binary data plus a labelling.  Binary is therefore the minimal
cardinality at which the \AgdaRecord{Distinction} record is
inhabitable in this framework, and higher arities reduce, in this
framework, to binary structure plus a labelling map.

\section{Normal Form}

An isomorphism of distinctions preserves the two boundary roles and has
inverse maps on carriers.  The normal-form theorem says that every
distinction is isomorphic, in this sense, to the canonical distinction
on \AgdaDatatype{Two}.

\begin{code}
record DistinctionIso (d e : Distinction) : Set₁ where
  private
    module D = Distinction d
    module E = Distinction e
  field
    to         : D.S → E.S
    from       : E.S → D.S
    to-left    : to D.left ≡ E.left
    to-right   : to D.right ≡ E.right
    from-left  : from E.left ≡ D.left
    from-right : from E.right ≡ D.right
    to-from    : (y : E.S) → to (from y) ≡ y
    from-to    : (x : D.S) → from (to x) ≡ x
\end{code}

The map to \AgdaDatatype{Two} is forced by the cover: an element is sent
to \AgdaInductiveConstructor{L} if it is the left boundary and to
\AgdaInductiveConstructor{R} if it is the right boundary.

\begin{code}
toTwo : (d : Distinction) → Distinction.S d → Two
toTwo d x with Distinction.cover d x
... | inj₁ _ = L
... | inj₂ _ = R

fromTwo : (d : Distinction) → Two → Distinction.S d
fromTwo d L = Distinction.left d
fromTwo d R = Distinction.right d

toTwo-left : (d : Distinction) → toTwo d (Distinction.left d) ≡ L
toTwo-left d with Distinction.cover d (Distinction.left d)
... | inj₁ _ = refl
... | inj₂ p = ⊥-elim (Distinction.separated d p)

toTwo-right : (d : Distinction) → toTwo d (Distinction.right d) ≡ R
toTwo-right d with Distinction.cover d (Distinction.right d)
... | inj₁ p = ⊥-elim (Distinction.separated d (sym p))
... | inj₂ _ = refl

toTwo-fromTwo : (d : Distinction) → (y : Two) → toTwo d (fromTwo d y) ≡ y
toTwo-fromTwo d L = toTwo-left d
toTwo-fromTwo d R = toTwo-right d

fromTwo-toTwo : (d : Distinction) → (x : Distinction.S d) → fromTwo d (toTwo d x) ≡ x
fromTwo-toTwo d x with Distinction.cover d x
... | inj₁ p = sym p
... | inj₂ p = sym p

two-normal-form : (d : Distinction) → DistinctionIso d Two-distinction
two-normal-form d = record
  { to         = toTwo d
  ; from       = fromTwo d
  ; to-left    = toTwo-left d
  ; to-right   = toTwo-right d
  ; from-left  = refl
  ; from-right = refl
  ; to-from    = toTwo-fromTwo d
  ; from-to    = fromTwo-toTwo d
  }
\end{code}

This is the whole normal-form argument.  No choice is made in the map:
the cover field gives only two possible cases, and the separation field
rules out the contradictory boundary cases.

\section{Endomorphisms of Two}

An endomorphism of \AgdaDatatype{Two} is fixed by its two values.  There
are therefore four cases: constant left, constant right, identity, and
swap.

\begin{code}
data EndoCase : Set where
  constL : EndoCase
  constR : EndoCase
  ident  : EndoCase
  swap   : EndoCase

interpret : EndoCase → Two → Two
interpret constL _ = L
interpret constR _ = R
interpret ident  x = x
interpret swap   L = R
interpret swap   R = L

Pointwise : (Two → Two) → (Two → Two) → Set
Pointwise f g = (x : Two) → f x ≡ g x
\end{code}

The classifier simply inspects the two values of a function.

\begin{code}
classify : (Two → Two) → EndoCase
classify f with f L | f R
... | L | L = constL
... | R | R = constR
... | L | R = ident
... | R | L = swap

classify-sound : (f : Two → Two) → Pointwise f (interpret (classify f))
classify-sound f L with f L | f R
... | L | L = refl
... | R | R = refl
... | L | R = refl
... | R | L = refl
classify-sound f R with f L | f R
... | L | L = refl
... | R | R = refl
... | L | R = refl
... | R | L = refl
\end{code}

Uniqueness follows by evaluating at \AgdaInductiveConstructor{L} and
\AgdaInductiveConstructor{R}.  Any attempted identification of two
different cases forces either \AgdaInductiveConstructor{L} to equal
\AgdaInductiveConstructor{R}, or conversely.

\begin{code}
case-unique : (c d : EndoCase) → Pointwise (interpret c) (interpret d) → c ≡ d
case-unique constL constL _ = refl
case-unique constL constR p = ⊥-elim (L≠R (p L))
case-unique constL ident  p = ⊥-elim (L≠R (p R))
case-unique constL swap   p = ⊥-elim (L≠R (p L))
case-unique constR constL p = ⊥-elim (R≠L (p L))
case-unique constR constR _ = refl
case-unique constR ident  p = ⊥-elim (R≠L (p L))
case-unique constR swap   p = ⊥-elim (R≠L (p R))
case-unique ident  constL p = ⊥-elim (R≠L (p R))
case-unique ident  constR p = ⊥-elim (L≠R (p L))
case-unique ident  ident  _ = refl
case-unique ident  swap   p = ⊥-elim (L≠R (p L))
case-unique swap   constL p = ⊥-elim (R≠L (p L))
case-unique swap   constR p = ⊥-elim (L≠R (p R))
case-unique swap   ident  p = ⊥-elim (R≠L (p L))
case-unique swap   swap   _ = refl

classify-unique :
  (f : Two → Two) → (c : EndoCase) → Pointwise f (interpret c) → classify f ≡ c
classify-unique f c p =
  case-unique (classify f) c
    (λ x → trans (sym (classify-sound f x)) (p x))
\end{code}

\section{Why the Closure is \texorpdfstring{$K_4$}{K4}}

Composition of endomorphisms is total: for any two endomorphism cases,
their composite is again an endomorphism of \AgdaDatatype{Two}, hence
is one of the four cases just classified.

\begin{code}
_∘₂_ : (Two → Two) → (Two → Two) → Two → Two
(f ∘₂ g) x = f (g x)

composeCase : EndoCase → EndoCase → EndoCase
composeCase c d = classify (interpret c ∘₂ interpret d)

endomorphisms-closed :
  (c d : EndoCase) → Σ EndoCase (λ e → Pointwise (interpret e) (interpret c ∘₂ interpret d))
endomorphisms-closed c d =
  composeCase c d , λ x → sym (classify-sound (interpret c ∘₂ interpret d) x)
\end{code}

The closure argument is then finite and rigid.  To represent the four
endomorphism cases as distinct and usable data, no additional structure
is assumed beyond what representation itself requires.  Unpacking this
requirement yields exactly the following three conditions.

\begin{enumerate}
\item \emph{Separation.} Distinct endomorphism cases must remain
  distinct on the carrier.  No identification of cases is allowed.
\item \emph{Realisation.} Every endomorphism case must be represented
  by a vertex of the carrier.
\item \emph{No surplus.} Every vertex of the carrier must arise from
  one of these cases.
\end{enumerate}

These conditions are the representation contract used in this note.
They make explicit what ``representation'' means here: each
endomorphism case is carried as a distinct and retrievable element of
the carrier, and every element of the carrier corresponds to such a
case.  Without separation, cases collapse; without realisation, cases
are missing; without the no-surplus condition, additional structure is
introduced.  The contract is argued and displayed; it is not disguised
as a theorem that precedes its own definition.

Under these three conditions, fewer than four vertices cannot represent
the four cases without identification.  If two cases are identified,
\AgdaFunction{case-unique} shows that the corresponding functions agree
pointwise; but the only way different rows of the four case table agree
is by forcing \AgdaInductiveConstructor{L} and
\AgdaInductiveConstructor{R} to coincide, contradicting
\AgdaFunction{L≠R}.  Thus separation and realisation together force at
least four vertices.

An additional vertex is not forced by the endomorphism data.  The
composite of any two represented cases is already one of the four
classified cases, by \AgdaFunction{endomorphisms-closed}.  A new vertex
would not be the value of any forced composition; it would violate no
surplus.

Finally, total composability is not another assumption.  Any ordered
pair of endomorphisms has a composite, so there is no missing
composability relation between distinct represented cases.  For the
underlying simple graph, this is exactly $K_4$.

The following small records package the three conditions.  The record
\AgdaRecord{FaithfulClosure} contains realisation and separation:
\AgdaField{realize} supplies a vertex for each case, and
\AgdaField{reflect-realize} prevents two cases from being identified.
The record \AgdaRecord{MinimalClosure} adds no surplus through
\AgdaField{no-extra}.

These conditions are therefore not independent choices once this notion
of representation has been fixed; they unpack that notion.

\begin{code}
record FaithfulClosure : Set₁ where
  field
    V               : Set
    realize         : EndoCase → V
    reflect         : V → EndoCase
    reflect-realize : (c : EndoCase) → reflect (realize c) ≡ c

record MinimalClosure : Set₁ where
  field
    closure  : FaithfulClosure
    no-extra : (v : FaithfulClosure.V closure) →
               Σ EndoCase (λ c → FaithfulClosure.realize closure c ≡ v)

K4Vertex : Set
K4Vertex = EndoCase

K4Edge : K4Vertex → K4Vertex → Set
K4Edge c d = ¬ (c ≡ d)

canonicalK4 : MinimalClosure
canonicalK4 = record
  { closure = record
      { V               = EndoCase
      ; realize         = λ c → c
      ; reflect         = λ c → c
      ; reflect-realize = λ c → refl
      }
  ; no-extra = λ v → v , refl
  }

realize-reflect :
  (m : MinimalClosure) →
  (v : FaithfulClosure.V (MinimalClosure.closure m)) →
  FaithfulClosure.realize (MinimalClosure.closure m)
    (FaithfulClosure.reflect (MinimalClosure.closure m) v)
  ≡ v
realize-reflect m v with MinimalClosure.no-extra m v
... | c , p = trans (sym (cong realize q)) p
  where
    fc = MinimalClosure.closure m
    realize = FaithfulClosure.realize fc
    reflect = FaithfulClosure.reflect fc
    reflect-realize = FaithfulClosure.reflect-realize fc

    q : c ≡ reflect v
    q = trans (sym (reflect-realize c)) (cong reflect p)
\end{code}

The three record fields are not three independent stipulations.  Each
one is exactly the load-bearing piece of one of the three
representation conditions, and the following lemmas name that load.
The proofs are short because the conditions are exactly the bijection
data; that is the point.

\emph{Separation is load-bearing.}  The field
\AgdaField{reflect-realize} forces \AgdaField{realize} to be injective.
If two cases were identified on the carrier, applying
\AgdaField{reflect} to the identification would identify the cases
themselves, contradicting the four-case enumeration of
\AgdaDatatype{EndoCase}.

\begin{code}
realize-injective :
  (m : MinimalClosure) →
  (c d : EndoCase) →
  FaithfulClosure.realize (MinimalClosure.closure m) c
    ≡ FaithfulClosure.realize (MinimalClosure.closure m) d →
  c ≡ d
realize-injective m c d p =
  trans (sym (FaithfulClosure.reflect-realize (MinimalClosure.closure m) c))
    (trans (cong (FaithfulClosure.reflect (MinimalClosure.closure m)) p)
      (FaithfulClosure.reflect-realize (MinimalClosure.closure m) d))
\end{code}

\emph{Realisation is load-bearing.}  The totality of
\AgdaField{realize} on \AgdaDatatype{EndoCase} is exactly the demand
that every endomorphism case is hosted by some vertex.  Without it,
some case has no representative and the carrier fails to support the
composition table at all.

\begin{code}
case-has-vertex :
  (m : MinimalClosure) →
  (c : EndoCase) →
  Σ (FaithfulClosure.V (MinimalClosure.closure m))
    (λ v → FaithfulClosure.realize (MinimalClosure.closure m) c ≡ v)
case-has-vertex m c =
  FaithfulClosure.realize (MinimalClosure.closure m) c , refl
\end{code}

\emph{No surplus is load-bearing.}  The field \AgdaField{no-extra}
forbids vertices that do not arise from any endomorphism case.  Without
it, the carrier may host data that is not forced by the endomorphism
set, breaking the claim that the closure adds nothing to its datum.

\begin{code}
vertex-from-case :
  (m : MinimalClosure) →
  (v : FaithfulClosure.V (MinimalClosure.closure m)) →
  Σ EndoCase
    (λ c → FaithfulClosure.realize (MinimalClosure.closure m) c ≡ v)
vertex-from-case m v = MinimalClosure.no-extra m v
\end{code}

Dropping \AgdaField{reflect-realize} re-admits identification of
cases.  Dropping the totality of \AgdaField{realize} re-admits cases
without vertices.  Dropping \AgdaField{no-extra} re-admits vertices
without cases.  In each case, what survives is no longer a
representation of the four-case set in the intended sense.  The three
fields are therefore not chosen because they yield $K_4$; within this
contract they are fixed because dropping any one breaks representation,
and $K_4$ follows.

The companion proves these three load-bearing lemmas against its own
slim records \AgdaRecord{FaithfulClosure} and \AgdaRecord{MinimalClosure}
so that the file remains self-contained and type-checks without the
larger source.  The full development \texttt{Void.lagda.tex} carries
the analogous statements against the richer \AgdaRecord{K4Record} used
there; the present file does not import them, it reproduces them at
the level of detail that this note needs.

This is the sense in which the closure yields $K_4$: the forced vertex
set is the four-case set, and the induced simple graph connects every
distinct pair.  The equations \AgdaField{reflect-realize} and
\AgdaFunction{realize-reflect} show that any carrier satisfying the
three conditions is equivalent to the four-case carrier.

Equivalently, for any \AgdaBound{m} : \AgdaRecord{MinimalClosure}, the
carrier
\[
\AgdaField{FaithfulClosure.V}\,(\AgdaField{MinimalClosure.closure}\,m)
\]
is in bijection with \AgdaDatatype{EndoCase} via \AgdaField{realize}
and \AgdaField{reflect}.  The field \AgdaField{reflect-realize}
proves one composite to be the identity on \AgdaDatatype{EndoCase};
\AgdaFunction{realize-reflect} proves the other composite to be the
identity on the carrier.  If composition on the carrier is read through
this bijection,
\[
v \star_m w =
\AgdaField{realize}_m
  (\AgdaFunction{composeCase}\,(\AgdaField{reflect}_m\,v)
                              (\AgdaField{reflect}_m\,w)),
\]
then reflecting the composite recovers the four-case composition table
of \AgdaDatatype{EndoCase}.  In this precise sense every minimal
closure satisfying separation, realisation, and no surplus is
isomorphic to the canonical four-case carrier.  Thus $K_4$ is not only
minimal; it is unique up to isomorphism under the stated constraints.

\section{Leibniz collapse, executed}

The earlier section on why two cannot occur without separation argued
that the principle is universal: it appears in classical logic as
substitutivity, in set theory as extensionality, in category theory as
Yoneda, in quantum mechanics as the indistinguishability that forces
the statistical structure.  Once a concrete carrier is chosen, the
same principle becomes a checkable term.  Martin-L\"of type theory
realises it in two lines: the induction principle for propositional
equality yields, by direct application, the Leibniz collapse.  If two
elements satisfy exactly the same predicates, they are propositionally
equal.

\begin{code}
leibniz-collapse :
  {S : Set} (a b : S) →
  ((P : S → Set) → P a → P b) →
  a ≡ b
leibniz-collapse a b ind = ind (λ x → a ≡ x) refl
\end{code}

The Agda term is one realisation of the principle, not its source.
The principle is older than the type checker and would survive any
particular choice of carrier; the type checker only confirms that the
realisation in this carrier introduces no hidden step.  Inside the
present setting, this means that no element of any
\AgdaRecord{Distinction} is silently a third position pretending to be
a fourth: the \AgdaField{separated} field already certifies that the
two boundary points cannot be Leibniz-collapsed onto each other, and
the \AgdaField{cover} field certifies that no further occurrence
escapes the two.

\section{Re-projection to the datum}

The $K_4$ closure does not replace the original distinction; it
presupposes it.  This section makes the dependency explicit at two
levels of strength so that no reader can mistake the closure for an
independent foundation, and so that the trivial direction is not
advertised as more than it is.

\emph{The structural content as a checked term.}  The bijection
between the carrier of any minimal closure and the four endomorphism
cases is not left to commentary.  It is bundled as a record and
constructed from the lemmas already proved: \AgdaField{reflect-realize}
is one round-trip, \AgdaFunction{realize-reflect} is the other.  This
is the load-bearing structural map.

\begin{code}
record Bijection (A B : Set) : Set where
  field
    to      : A → B
    from    : B → A
    to-from : (b : B) → to (from b) ≡ b
    from-to : (a : A) → from (to a) ≡ a

closure-bijection-EndoCase :
  (m : MinimalClosure) →
  Bijection (FaithfulClosure.V (MinimalClosure.closure m)) EndoCase
closure-bijection-EndoCase m = record
  { to      = FaithfulClosure.reflect         (MinimalClosure.closure m)
  ; from    = FaithfulClosure.realize         (MinimalClosure.closure m)
  ; to-from = FaithfulClosure.reflect-realize (MinimalClosure.closure m)
  ; from-to = realize-reflect m
  }
\end{code}

This term is the actual dependency.  Every minimal closure has a
carrier in checked bijection with \AgdaDatatype{EndoCase}; by
\AgdaFunction{two-normal-form}, \AgdaDatatype{EndoCase} is built from
\AgdaDatatype{Two}, the canonical carrier of any binary distinction.
The four-vertex structure is therefore not an independent foundation;
it is, up to the bijection just constructed, the endomorphism set of
the normal form of any binary distinction.

\emph{The constant.}  At the bare level of returning a value of type
\AgdaRecord{Distinction}, every minimal closure determines a
distinction, namely the canonical one.  The corresponding Agda term is
a constant function.  The reason it is a constant and not a richer
structural map is honest: a \AgdaRecord{Distinction} has a two-element
carrier, while the closure carrier has four; no map of records
\AgdaBound{m} : \AgdaRecord{MinimalClosure} \AgdaSymbol{→}
\AgdaRecord{Distinction} can be a bijection on carriers.  The actual
bijective content lives one level below, in
\AgdaFunction{closure-bijection-EndoCase} above.  The constant below
is the bookkeeping shadow of that bijective content, recorded as such.

\begin{code}
k4-presupposes-distinction : MinimalClosure → Distinction
k4-presupposes-distinction _ = Two-distinction
\end{code}

The full formalisation uses a richer \AgdaRecord{K4Record}.  The
structural dependency carries over without change: the closure carrier
is not an independent foundation, it is, by
\AgdaFunction{closure-bijection-EndoCase}, in checked bijection with
the endomorphism set of the binary normal form.

\part*{Part III.\quad Frame}

\section{Scope Clarification}

The scope of this note is narrow by design.

\emph{What is not claimed.} Nothing in the preceding sections asserts
that a binary distinction takes place. The development never builds
an inhabitant of \AgdaRecord{Distinction} without supplied data; every
Distinction-valued construction below the abstract takes a Distinction
as input. The proofs are silent on the question of whether such an
input exists in any external sense.

\emph{What is conceptually argued, not formally proved.} The
introductory paragraphs argue that any formalisation at all already
presupposes a field of separability: relations, morphisms, judgements,
and equations all require distinguishable positions before they can be
used.  This is a remark on the conditions under which formalisation is
possible, not a theorem.  Nothing in the Agda code attempts to prove it,
and the formal results below do not depend on it: they begin once the
data are granted.

\emph{What is machine-proved.} The conditional consequence is what
the Agda code establishes. Given a structure with two
distinguishable, exhaustive points, the normal form is
\AgdaDatatype{Two}; given the endomorphisms of that normal form,
representing the four cases with separation, realisation, and no
surplus yields the four-case carrier with complete simple adjacency;
and the closure entails the data from which it was built. The result
therefore does not say that every theory begins with such a distinction,
nor that every mathematical object reduces to \AgdaDatatype{Two}. It
says only what follows when the stated data are granted.

No further constructions are treated here. They are outside the
present claim.

\section{Placement}

The note is brief enough to mistake for naivety; this section locates
it in three lineages so that the contribution is read for what it is.

\emph{Spencer-Brown.}  The thesis that distinction is the foundational
act is at least as old as \emph{Laws of Form} (1969).  The present
note neither defends nor extends that thesis.  It records what holds
once a distinction is granted, in a form a type checker can verify.
The contribution is the precision of the conditional, not the
originality of the antecedent.

\emph{Lawvere.}  ``Adjointness in Foundations'' (1969) treats truth
and falsity as adjoint to the act of distinguishing.  The present note
operates one level below: before truth-values are assigned, before
propositions are evaluated.  \AgdaDatatype{Two} is a fresh carrier,
deliberately unread as $\{\bot,\top\}$, so that the closure result
depends on separation alone and not on a prior reading of the two
points as truth-values.

\emph{Univalent foundations.}  \AgdaDatatype{Two} is an h-set, and the
development uses \texttt{--without-K}.  No higher-dimensional
structure is assumed or invoked.  The note is therefore compatible
with both intensional Martin-Löf type theory and a univalent reading;
nothing in the closure result depends on the choice of identity
discipline.

What is offered here is not a competing foundation, and not an
argument that distinction-based foundations are superior to others.
It is one machine-checked statement at a place where earlier work has
been mostly informal: that the closure of binary distinction is $K_4$,
unique up to isomorphism, under three precisely fixed conditions of
representation, and that this closure points back to the data from
which it was built.

\section{Outlook}

The core result stops at the framework-forced closure of binary distinction.  Its
role in the larger development is not to prove everything at once, but
to fix the first carrier on which later constructions may be tested.
Once the four-case closure is fixed, further claims can be separated
cleanly: some are formal consequences of the core, some are additional
constructions over it, and some belong only to the interpretive layer.

The next formal step in \emph{Void and Form} is therefore not a jump to
a global conclusion.  It is the controlled question of what can be built
on the surviving carrier without adding new generators.  This note only
isolates the starting point.

\subsection*{Interpretive continuation}

The later volume \emph{Form} belongs here, at the end of the companion,
not at the beginning.  It is not the kernel and it is not evidence for
the kernel.  The direction is one-way: the checked kernel exports an
invariant ledger; \emph{Form} reads that ledger structurally and, where
the reading is explicit enough, physically.  A failed physical reading
would not refute the normal-form theorem or the four-case closure.  It
would refute that interpretation.

This does not make \emph{Form} decorative.  Its empirical surface is
deliberately exposed.  The physical identifications are tested by
correction chains, sigma deviations, and failure conditions.  In the
current ledger, the post-correction comparisons reported in
\texttt{Form.lagda.tex} are:

\begin{center}
\begin{tabular}{lp{0.66\textwidth}}
\hline
Observable & Reported comparison in \emph{Form} \\
\hline
$\alpha^{-1}$ & second-order correction
$137 + 11/306 + 295/(306\cdot137^2)$, within $0.4\sigma$ of CODATA; the
alternative interpolation degrees are reported as $2445\sigma$ and
$4891\sigma$ failures. \\
$m_p/m_e$ & correction chain $949\sigma \to 20\sigma \to 0.08\sigma$. \\
$m_\mu/m_e$ & correction chain $36\sigma \to 13\sigma \to 0.12\sigma$. \\
$m_\tau/m_\mu$ & remains closed at about $0.1\sigma$. \\
$\sin^2\!\theta_W$ & weak-mixing chain
$7.7\sigma \to 5.0\sigma \to 3.5\sigma \to 0.001\sigma$. \\
$\Omega_m$ & $K_4$ gives $\Omega_m \approx 0.3183$, within
$0.5\sigma$ of Planck; the $K_3$ and $K_5$ alternatives are reported as
$10.9\sigma$ and $11.8\sigma$ away. \\
\hline
\end{tabular}
\end{center}

The same separation governs the open tests.  The B-meson
lepton-universality anomaly is read through the muon bridge factor
$16/69 \approx 0.232$ as a Wilson-coefficient target.  The doubly
charmed baryon section reads the observed $\Xi_{cc}$ pair as occupation
of a forced two-slot doublet ledger, not as a mass prediction.  The
frontier register names exactly two gravitational-wave polarizations, a
neutrino seesaw at loop depth $5$, Planck-scale curvature $R = 12$, no
fourth generation, and no extra spacetime dimension.  These are
interpretive and empirical claims.  They are where the larger project
becomes vulnerable to physics.

\section{Navigation into the book}

The corresponding self-contained development is in the repository
\href{https://github.com/de-johannes/Void-and-Form}{de-johannes/Void-and-Form}.
The main file is \texttt{Void.lagda.tex}.  The route through the larger
book is:

\begin{itemize}
\item \emph{Void as boundary}: \emph{What Void Does Not Name},
  \emph{Why Formal Systems Cannot Begin Themselves}, and
  \emph{Representation Requires A Cut}.
\item \emph{The binary threshold}: \emph{Why The First Cut Is Binary},
  \emph{The Formal Threshold}, and the record \texttt{Distinction}.
\item \emph{Normal form and classification}: \emph{Elimination and
  Classification}, \emph{Two Normal Form}, and the endomorphism
  classification of \texttt{Two -> Two}.
\item \emph{Four-vertex closure}: \emph{Graph Presentation},
  \emph{The K4 Invariant Record}, \emph{K4 Representation Theory},
  and \texttt{K4Record-is-canonical}.
\item \emph{Return to the datum}: \texttt{record-presupposes-distinction},
  the formal witness that the surviving closure entails the originating
  distinction.
\item \emph{Interpretive continuation}: the later volume \emph{Form},
  where origin, contact, spatial separation, temporal order, physical
  reading, contingency, and induction are treated as structural
  interpretations rather than new generators.
\end{itemize}

\end{document}